---
title: "H_Index_Rcpp"
author: "Nathan Quan"
date: "3/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/nathansquan/Documents/DataScience/Segregation/')
```

# Overview

## Goal

Speed up data ingest and calculation of H Index using the `Rcpp` package.

First we will demonstrate how to to calculate H Indices for a list of MSAs using R and benchmark it. Then we will implement C++ code to speed up these calculations and data ingest and compare the time it takes between the two methods.

At a high level, both methods require us to:

1. Get tract-level ACS income data for each MSA
2. Calculate H Indices

## Load Packages
```{r load tidyverse echo=FALSE}
library(tidyverse)
```


# Data ingest and initial cleaning

## MSA Data

MSA data are from the Census Bureau. I use the 2017 definitions of MSAs. This data contain the names and FIPS codes for all counties and the MSA in which they belong.

https://www.census.gov/geographies/reference-files/time-series/demo/metro-micro/delineation-files.html

```{r import MSA data}
msa_2017 <- readxl::read_excel("data/raw/CBSA_Aug_2017.xls")
#save(msa_2017, file = "MSA_2017.RData")
```


## Cleaning MSA data

* Drop Notes at bottom of data frame (last 4 lines)

* Join State FIPS with County FIPS

* Filter missing and only to MSAs

```{r Clean MSA data}

str(msa_2017)

msa_2017 <- msa_2017 %>%
  head(-4)

msa_2017$FIPS64 <- str_c(msa_2017$`FIPS State Code`, msa_2017$`FIPS County Code`)

msa_2017 <- msa_2017 %>%
  filter(!is.na(`CBSA Title`)) %>%
  filter(`Metropolitan/Micropolitan Statistical Area` == "Metropolitan Statistical Area") %>%
  select(`CBSA Code`, `CBSA Title`, `Metropolitan/Micropolitan Statistical Area`, `County/County Equivalent`, FIPS64)

```


## Load ACS variables

Here we load a list of all variables for the 2017 ACS 5-year survey using the `tidycensus` package. Then we define a list of the income variables from ACS.
```{r}
vv <- tidycensus::load_variables(2017, 'acs5', cache = T)

###Define 5-year ACS variables of interest

## INCOME VARIABLE

# Define income vars
vars_inc <- c(

  ## TABLE B19001: Household Income categories - TOTAL POP
  "B19001_002", # Income <10k
  "B19001_003", # Income 10-15k
  "B19001_004", # Income 15-20k
  "B19001_005", # Income 20-25k
  "B19001_006", # Income 25-30k
  "B19001_007", # Income 30-35k
  "B19001_008", # Income 35-40k
  "B19001_009", # Income 40-45k
  "B19001_010", # Income 45-50k
  "B19001_011", # Income 50-60k
  "B19001_012", # Income 60-75k
  "B19001_013", # Income 75-100k
  "B19001_014", # Income 100-125k
  "B19001_015", # Income 125-150k
  "B19001_016", # Income 150-200k
  "B19001_017" # Income 200+

)
```



# The R-only way


# Create a List of MSA objects

Now that we have a list of all counties that are part of MSAs in the U.S., we need to place each county's FIPS code into an object for each MSA.

The MSA objects will be named by their cbsa code.

The for loop iterates over the CBSA code variable to create a list element named using the CBSA code in the current iteration. From there, we use `dplyr::filter()` to find all rows in the `msa_2017` data frame that has the current CBSA code.

Finally, we check the number of unique CBSA codes in `msa_2017` to make sure the for loop works. There are 390 unique CBSA codes in `msa_2017` and 390 elements in `MSA_inc`.

```{r}
MSA_inc <- lst() #create empty list

for (cbsa_code in msa_2017$`CBSA Code`) {
  MSA_inc[[cbsa_code]] <- msa_2017 %>%
    filter(`CBSA Code` %in% cbsa_code)
}

# Check
length(unique(as.numeric(msa_2017$`CBSA Code`)))
```

